A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2023-07-13 paper/fascd.tex
address more PF comments

This paper is organized as follows.  In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, with several example problems.  Section \ref{sec:cd} reviews the basics of CDs, as proposed by Tai \cite{Tai2003}, and their induced iterations.  Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} presents defect constraints generated via monotone restriction operators \cite{GraeserKornhuber2009}.  Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD.  Section \ref{sec:vcycle} then states the FAScd V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.  Section \ref{sec:results} gives FAScd numerical results from a parallel Firedrake \cite{Rathgeberetal2016} implementation for the problems introduced in Section \ref{sec:vi}.  The work concludes with discussion and outlook in Section \ref{sec:discussion}, and an appendix on how FAScd reduces to classical FAS multigrid for PDEs.

@@

\begin{table}[H]
\begin{tabular}{llc}
\toprule
\emph{name}  & \emph{action}  & \emph{linear?} \\ \midrule
canonical prolongation        & $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ & \,\checkmark \\
canonical (dual) restriction  & $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ & \,\checkmark \\
(nodal) injection             & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
maximum restriction           & $\maxR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
minimum restriction           & $\minR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
\bottomrule
\end{tabular}
\caption{Transfer operators used in this paper.}
\label{tab:transfers}
\end{table}

@@

\pef{I don't like the name ``monotone restrictions''. The word restriction is strongly associated with operating on the dual space. I would prefer to call them ``monotone injections'', since that is the matching operator.}
The monotone restrictions $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices). These operators are constructed to enforce the property
\begin{equation}
\minR z \le z \le \maxR z,  \label{eq:fe:monotonerestrictionprops}
\end{equation}
which does not hold for the standard injection operator $\iR$.
To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.  For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
\begin{align}
(\maxR z)(x_p^{j-1}) &= \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \label{eq:fe:monotonerestrictions} \\
(\minR z)(x_p^{j-1}) &= \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}. \notag
\end{align}
For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine vertices $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotonerestrictionprops} follows.
On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
\pef{It would be good to have a diagram here. I could write one in tikz, if you don't have one handy.}

@@

Descending coarser levels, $j=J-1,\dots,1,0$, we apply the monotone restrictions to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
\begin{equation}
\underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \qquad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}. \label{eq:fe:chilevels}
\end{equation}
Observe that
\begin{equation}
\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J. \label{eq:fe:chiordering}
\end{equation}
A unilateral $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.  The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; similar comments apply to $\overline{\chi}^{j}$.

%REGENERATE using https://github.com/bueler/mg-glaciers:
%$ cd mg-glaciers/py/
%$ [modify decomposition() inside visualize.py to turn off axes and print \gamma for obstacle]
%$ ./obstacle.py -J 5 -jcoarse 1 -random -randommodes 8 -diagnostics -o defect.pdf
%fine level 5 (m=63): using 20 V(1,0) cycles -> 38.750 WU
\begin{figure}[ht]
\includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
\caption{Unilateral LDCs $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum restriction: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
\label{fig:chiphilevels}
\end{figure}

@@

The V-cycle algorithm in this section applies smoothers in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.  It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}.  In fact, it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).

Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ by incorporating information from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$,\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.} and are constructed so that the corrected iterate remains admissible.  Proceeding inductively, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j+1$ \pef{I changed the sign here from $j-1$ to $j+1$, double-check} are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  At the bottom the coarsest-level VI is then solved, resulting in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way (Algorithm \ref{alg:basiccd-tele}), up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.  Then the next admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.

@@

As we descend we define new $j$th-level iterates
\begin{equation}
w^j = \iR(w^{j+1} + y^{j+1}).  \label{eq:fe:definew}
\end{equation}
(These are the states before smoothing or correction.)  From $w^j$ we define $\ell^j \in (\cV^j)'$:
\begin{equation}
\ell^j = \begin{cases} \ell^J, & j=J \\
                       f^j(w^j) + R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
\end{equation}
Observe that injection $\iR$ coarsens the (primal) iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens (dual) linear functionals; recall Table \ref{tab:transfers}.

On descent, at levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
\begin{equation}
\ip{f^j(w^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j, \label{eq:fe:downvi}
\end{equation}
for which the initial iterate is $y_{(0)}^j=0$.  (Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved but for $z^j \in \mathcal{U}^j$:
\begin{equation}
\ip{f^j(w^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
\end{equation}
The nontrivial initial iterate for this problem is found by prolonging $z^{j-1}$ and using it to correct the current (i.e.~down-smoothed) correction on the level:
\begin{equation}
z_{(0)}^j = y^j + P z^{j-1}.  \label{eq:fe:upwardinitial}
\end{equation}
An exception is that $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible by \eqref{eq:fe:upwardsuminclusion}.

@@

Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.\footnote{We could write ``$V(\id{down},\id{up})$'', but we prefer to emphasize the number of subsets used on each level rather than the number of smoother sweeps in a given set. \pef{I don't understand what the number of subsets used on each level means.}}  If \id{up} $=0$ then it becomes a $V(1,0)$ cycle and if \id{down} $=0$ then it is a $V(0,1)$ cycle; in the former case it generalizes Algorithm 4.7 in \cite{GraeserKornhuber2009} by allowing nonlinear operators $f$, any smoother, and applying FAS-type corrections.  In unilateral cases, one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}; in this sense $V(1,1)=V(1,0)+V(0,1)$.

@@

In this section we consider the implementation of FAScd methods.  We address convergence criteria, the F-cycle extension, and an active set Newton method used as a smoother and coarse solver.

Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.  To construct appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}.  Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding basis function.  Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.  If $m_D$ is the number of vertices $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.  Finally, write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.  Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
\begin{align}
\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
\end{align}
for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
\pef{I wonder if this definition will confuse people, because they won't know how to plug in $\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ into your definition of $\perp$ ($\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ is not in $\RR^n$). Maybe it would be easier to just use the spelled-out version below, and not use this notation?}

By definition, MCP \eqref{eq:fe:mcp} says that certain node-wise conditions hold for $u^J$ at $x_p \in \mathcal{T}^J$.  Let $r(u^J)_p = \ip{f^J(w^J)-\ell^J}{\psi_p}$ be the pointwise (ordinary) residual.  If $x_p \notin \partial_D\Omega$ then one of these four conditions hold:
\begin{itemize}
\item $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $r(u^J)_p = 0$,
\item $\underline{\gamma}^J(x_p)=u^J(x_p)<\overline{\gamma}^J(x_p)$ and $r(u^J)_p \ge 0$,
\item $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $r(u^J)_p \le 0$,
\item $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$.
\end{itemize}
Otherwise, if $x_p \in \partial_D\Omega$ then
\begin{itemize}
\item $u^J(x_p)=g_D^J(x_p)$ and $\lambda_p=r(u^J)_p$.
\end{itemize}
The cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched} (\emph{both active}), and \emph{boundary}, respectively.  Where \emph{pinched} applies, $r(u^J)_p$ can have any real value.

Thus \eqref{eq:fe:mcp} allows us to construct expressions whose norms should go below tolerance.  First, interpreting the MCP directly, for $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} $\rNC(w^J)$ from the five cases:
\begin{equation}
\rNC(w^J)_p = \begin{cases}
    r(w^J)_p, & \underline{\gamma}^J(x_p) < w^J(x_p) < \overline{\gamma}^J(x_p), \\
    \min\{r(w^J)_p,0\}, & w^J(x_p) = \underline{\gamma}^J(x_p), \\
    \max\{r(w^J)_p,0\}, & w^J(x_p) = \overline{\gamma}^J(x_p), \\
    0, & \underline{\gamma}^J(x_p)=w^J(x_p)=\overline{\gamma}^J(x_p), \\
    w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega. \end{cases} \label{eq:rNC}
\end{equation}
The first four cases assume $x_p \notin \partial_D\Omega$.  For PDEs one uses only the first and last cases.

In summary, $\rNC : \mathcal{K}^J \to \RR^{m_J}$ is a computable nonlinear map defined using the discretization $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$.  A value $\rNC(w^J)_p$ is nonzero only at vertices where \eqref{eq:fe:mcp} is violated, either because $r(w^J)_p$ remains nonzero where the constraints are inactive, or because it has the wrong sign at an active node, or because the boundary condition is not satisfied.

\pef{I think this is all fairly standard and we could abbreviate this section substantially.}

Direct implementation of \eqref{eq:rNC} is, however, sensitive to rounding error because $\rNC$ is a discontinuous function of $w^J(x_p)$.  For example, if $w^J(x_p)$ exceeds $\underline{\gamma}^J(x_p)$ by $10^{-16}$, then, instead of returning the minimum of $r(w^J)_p$ and zero (the second case), the first case applies, so $\rNC(w^J)_p$ might be large and positive.  This situation would falsely suggest that the iteration is far from converging.  On the other hand, formula \eqref{eq:rNC} applies as stated if $\underline{\gamma}^J(x_p)=-\infty$ or $\overline{\gamma}^J(x_p)=+\infty$.  Compare \eqref{eq:rSS} in this case, wherein the difference $w^J(x_p) - \underline{\gamma}^J(x_p)$ is computed.

>>>>

This paper is organized as follows.  In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, with several example problems.  Section \ref{sec:cd} reviews the basics of CDs, as proposed by Tai \cite{Tai2003}, and their induced iterations.  Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} presents defect constraints generated via monotone injection operators \cite{GraeserKornhuber2009}.  Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD.  Section \ref{sec:vcycle} then states the FAScd V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.  Section \ref{sec:results} gives FAScd numerical results from a parallel Firedrake \cite{Rathgeberetal2016} implementation for the problems introduced in Section \ref{sec:vi}.  The work concludes with discussion and outlook in Section \ref{sec:discussion}, and an appendix on how FAScd reduces to classical FAS multigrid for PDEs.

@@

\begin{table}[H]
\begin{tabular}{llc}
\toprule
\emph{name}  & \emph{action}  & \emph{linear?} \\ \midrule
canonical prolongation        & $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ & \,\checkmark \\
canonical (dual) restriction  & $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ & \,\checkmark \\
(nodal) injection             & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
maximum injection           & $\maxR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
minimum injection           & $\minR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
\bottomrule
\end{tabular}
\caption{Transfer operators used in this paper.}
\label{tab:transfers}
\end{table}

@@

%\pef{I don't like the name ``monotone restrictions''. The word restriction is strongly associated with operating on the dual space. I would prefer to call them ``monotone injections'', since that is the matching operator.}
% Agreed.  "monotone restriction" is used in GraeserKornhuber2009, but happy not to follow.

The monotone injections $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices). These operators are constructed to enforce the property
\begin{equation}
\minR z \le z \le \maxR z,  \label{eq:fe:monotonerestrictionprops}
\end{equation}
which does not hold for the standard injection operator $\iR$.
To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.  For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
\begin{align}
(\maxR z)(x_p^{j-1}) &= \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \label{eq:fe:monotonerestrictions} \\
(\minR z)(x_p^{j-1}) &= \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}. \notag
\end{align}
For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine vertices $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotonerestrictionprops} follows.
On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
\pef{It would be good to have a diagram here. I could write one in tikz, if you don't have one handy.}

@@

Descending coarser levels, $j=J-1,\dots,1,0$, we apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
\begin{equation}
\underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \qquad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}. \label{eq:fe:chilevels}
\end{equation}
Observe that
\begin{equation}
\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J. \label{eq:fe:chiordering}
\end{equation}
A unilateral $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.  The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; similar comments apply to $\overline{\chi}^{j}$.

%REGENERATE using https://github.com/bueler/mg-glaciers:
%$ cd mg-glaciers/py/
%$ [modify decomposition() inside visualize.py to turn off axes and print \gamma for obstacle]
%$ ./obstacle.py -J 5 -jcoarse 1 -random -randommodes 8 -diagnostics -o defect.pdf
%fine level 5 (m=63): using 20 V(1,0) cycles -> 38.750 WU
\begin{figure}[ht]
\includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
\caption{Unilateral LDCs $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum injection: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
\label{fig:chiphilevels}
\end{figure}

@@

The V-cycle applies smoothers in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.  It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).

Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ by incorporating information from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.}  The corrections are constructed, inductively, so that iterates remain admissible at every level.  Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  Solving the coarsest-level VI results in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way (Algorithm \ref{alg:basiccd-tele}), up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.  Then the next finest-level, admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.

@@

As we descend we define new $j$th-level iterates
\begin{equation}
w^j = \iR(w^{j+1} + y^{j+1}).  \label{eq:fe:definew}
\end{equation}
(These are the states before smoothing or correction.)  From $w^j$ we define $\ell^j \in (\cV^j)'$:
\begin{equation}
\ell^j = \begin{cases} \ell^J, & j=J \\
                       f^j(w^j) + R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
\end{equation}
Observe that injection $\iR$ coarsens the (primal) iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens (dual) linear functionals; recall Table \ref{tab:transfers}.  At levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
\begin{equation}
\ip{f^j(w^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j, \label{eq:fe:downvi}
\end{equation}
for which the initial iterate is $y_{(0)}^j=0$.  (Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)

On ascent, $j=0,1,\dots,J$, the same VI is solved but for $z^j \in \mathcal{U}^j$:
\begin{equation}
\ip{f^j(w^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
\end{equation}
The nontrivial initial iterate for this problem is found by prolonging $z^{j-1}$ and using it to correct the current (i.e.~down-smoothed) correction on the level:
\begin{equation}
z_{(0)}^j = y^j + P z^{j-1}.  \label{eq:fe:upwardinitial}
\end{equation}
An exception is that $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible by \eqref{eq:fe:upwardsuminclusion}.

@@

Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle with \id{down}, \id{up} smoother iterations on each level.  For \id{up} $=0$ it becomes a $V(1,0)$ cycle, and for \id{down} $=0$ a $V(0,1)$ cycle.  The $V(1,0)$ cycle is similar to Algorithm 4.7 in \cite{GraeserKornhuber2009}, but generalizes it by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.  One iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}) over CD\footnote{Strictly speaking, this statement is restricted to unilateral cases.} \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Algorithm \ref{alg:basiccd-tele}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.

@@

Implementations of FAScd methods must choose a concrete convergence criterion.  In this section we also construct the F-cycle extension, and we describe the active set Newton method we will use as a smoother and coarse solver.

%Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding basis function.  Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.  If $m_D$ is the number of vertices $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.  Finally, write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.  Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
%\begin{align}
%\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
%I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
%\end{align}
%for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
%\pef{I wonder if this definition will confuse people, because they won't know how to plug in $\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ into your definition of $\perp$ ($\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ is not in $\RR^n$). Maybe it would be easier to just use the spelled-out version below, and not use this notation?}
% Agreed.  Greatly simplified

Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.  To construct appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}.  By definition, this says that certain node-wise conditions hold for $u^J$ at $x_p \in \mathcal{T}^J$.  Let $r(u^J)_p = \ip{f^J(w^J)-\ell^J}{\psi_p}$ be the pointwise (ordinary) residual.  If $x_p \notin \partial_D\Omega$ then one of these four conditions hold:
\begin{itemize}
\item $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $r(u^J)_p = 0$,
\item $\underline{\gamma}^J(x_p)=u^J(x_p)<\overline{\gamma}^J(x_p)$ and $r(u^J)_p \ge 0$,
\item $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $r(u^J)_p \le 0$,
\item $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$.
\end{itemize}
Otherwise, if $x_p \in \partial_D\Omega$ then
\begin{itemize}
\item $u^J(x_p)=g_D^J(x_p)$.
\end{itemize}
These five cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched} (\emph{both active}), and \emph{boundary}, respectively.  Where \emph{pinched} or \emph{boundary} apply, $r(u^J)_p$ can have any value.

This MCP interpretation allows us to construct expressions whose norms should go below tolerance.  For $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} $\rNC(w^J)$ from the five cases:
\begin{equation}
\rNC(w^J)_p = \begin{cases}
    r(w^J)_p, & \underline{\gamma}^J(x_p) < w^J(x_p) < \overline{\gamma}^J(x_p), \\
    \min\{r(w^J)_p,0\}, & w^J(x_p) = \underline{\gamma}^J(x_p), \\
    \max\{r(w^J)_p,0\}, & w^J(x_p) = \overline{\gamma}^J(x_p), \\
    0, & \underline{\gamma}^J(x_p)=w^J(x_p)=\overline{\gamma}^J(x_p), \\
    w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega. \end{cases} \label{eq:rNC}
\end{equation}
The first four cases assume $x_p \notin \partial_D\Omega$.  For PDEs one uses only the first and last cases.

%\pef{I think this is all fairly standard and we could abbreviate this section substantially.}
% Yes.
